---
title: "Moderated mediation and mediated moderation using rmedsem"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE, message = FALSE,
  comment = "#>"
)
```

It is often of interest to assess whether a mediation effect is invariant across contexts, or
whether its strength depends on the level of another variable. In such cases, the indirect effect of
an independent variable on an outcome via a mediator is moderated by a third variable (the moderator). 
These types of effects are often referred to as moderated mediation or mediated
moderation effects (Preacher et al., 2007). 

Currently, `rmedsem` implements moderated mediation and mediated moderation.
Mediated moderation is available for cases
corresponding to "Model 2" from Preacher et al., 2007 shown in the following graph:


```{r, echo=FALSE}
#| echo: false
#| warning: false
#| eval: true
m <- '
M ~ X + W + XW
Y ~ X + M + W + XW
'

plot_m <- semPlot::semPlotModel_lavaanModel(m)

semPlot::semPaths(plot_m,
                  rotation=2,
                  sizeMan = 5,
                  sizeMan2 = 3,
                  sizeLat = 12, 
                  layout = "tree3",
                  style = "mx",
                  residuals = FALSE,
                  edge.label.cex = 0.6,
                  nCharNodes = 10)
```

Mathematically, this model can be specified as 

\begin{align*}
M &= a_0 + a_1 X + a_2 W + a_3 (X \times W) + \varepsilon_M\\
Y &= b_0 + b_1 M + b_2 X + b_3 (X \times W) + \varepsilon_Y
\end{align*}

where $X$ is the independent variable, $Y$ is the dependent variable, $M$ is the mediator, and $W$ is the moderator. The interaction term $X \times W$ is included in the equations for $M$ and $Y$.

To estimate moderated mediation and mediated moderation with `rmedsem`, we need to specify the model using the [`modsem`](https://modsem.org) package which allows to estimate structural equation models with interaction terms using `lavaan`. We specify and estimate the following model (corresponding to the conceptual model above) using the dataset `rmedsem::mchoice`:


```{r}
library(modsem)

model <- "
  OwnLook =~ smv_attr_face + smv_attr_body + smv_sexy
  SelfEst =~ ses_satis + ses_qualities + ses_able_todo
  MentWell =~ mwb_optimistic + mwb_useful + mwb_energy
  OwnPers =~ smv_kind + smv_caring + smv_understanding +
    smv_make_laughh + smv_funny + smv_sociable

  MentWell ~ OwnLook + SelfEst + OwnPers + OwnPers:OwnLook
  SelfEst ~ OwnLook + OwnPers + OwnPers:OwnLook
"

est <- modsem(model, data = rmedsem::mchoice, method="lms")
```

## Mediated Moderation
*Mediated moderation* treats the interaction term `X:W` as the independent variable and asks whether its effect on `Y` is carried through `M`.
In our example, we can test whether the interaction between `OwnLook` and `OwnPers` on `MentWell` is mediated by `SelfEst`.

```{r}
library(rmedsem)
rmedsem(indep="OwnPers:OwnLook", dep="MentWell", med="SelfEst", mod=est)
```


## Moderated Mediation

*Moderated mediation* asks whether the indirect effect of an exposure `X` on outcome `Y` via mediator `M` varies across levels of a third variable `W`.  
Using the model from the previous example, we test how the indirect path from `OwnLook` to `MentWell` through `SelfEst` depends on `OwnPers`.

```{r}
rmedsem(indep="OwnLook", dep="MentWell", med="SelfEst", mod=est,
        moderator="OwnPers")
```

In this case the difference between a *moderated mediation* and *mediated moderation* is
purely semantic. Indeed, the indirect and total moderation effect when interpreted as 
a moderated mediation is the exact same as the indirect and total effect 
in the previous example ($.07$ and $.08$).

That being said, moderated mediations can be more complex in nature than mediated moderations, 
where the moderating variable `W` can affect the paths of model differently. Here 
we can for example see a moderated mediation where `OwnPers` not only affects 
the path from `OwnLook` to `SelfEst` and `OwnLook` to `MentWell`, but also 
the path from `SelfEst` to `MentWell`.

```{r, warning=FALSE}
model2 <- "
  OwnLook =~ smv_attr_face + smv_attr_body + smv_sexy
  SelfEst =~ ses_satis + ses_qualities + ses_able_todo
  MentWell =~ mwb_optimistic + mwb_useful + mwb_energy
  OwnPers =~ smv_kind + smv_caring + smv_understanding +
    smv_make_laughh + smv_funny + smv_sociable

  SelfEst ~ OwnLook + OwnPers + OwnPers:OwnLook
  MentWell ~ OwnLook + SelfEst + OwnPers + OwnPers:OwnLook + OwnPers:SelfEst
"

est2 <- modsem(model2, data = rmedsem::mchoice, method="lms")
```

`rmedsem()` will automatically detect the paths wich are moderated by the `moderator`
and tailor the output accordingly.


```{r}
rmedsem(indep="OwnLook", dep="MentWell", med="SelfEst", mod=est2,
        moderator="OwnPers")
```
