---
title: "Working with the results"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE, message = FALSE,
  comment = "#>"
)
set.seed(1)
```

`rmedsem()` returns an object of class `rmedsem`. Besides printing it, there
are several functions to summarize the results, extract estimates for further
processing, and visualize them. We use the simple mediation model from the
[Getting started](../index.html#getting-started) section as an example:

```{r setup}
library(rmedsem)

mod.txt <- "
  read ~ math
  science ~ read + math
"
mod <- lavaan::sem(mod.txt, data = hsbdemo)
out <- rmedsem(mod, indep = "math", med = "read", dep = "science")
```

## Detailed output and compact summary

Printing the object gives a detailed, step-by-step description of the tests
of the indirect effect and of the Baron & Kenny and Zhao, Lynch & Chen
approaches:

```{r}
out
```

`summary()` condenses the same results into a table of the indirect (for each
estimation method), direct and total effects, the resulting type of mediation
and the effect sizes:

```{r}
s <- summary(out)
s
```

The summary is itself an object whose elements can be used in further
analyses, e.g., the type of mediation or the table of effects:

```{r}
s$mediation
s$effects
```

## Extracting estimates

The standard extractor functions `coef()`, `confint()` and `nobs()` work for
`rmedsem` objects:

```{r}
coef(out)
confint(out)
nobs(out)
```

The indirect effect is estimated with several methods (here Sobel, Delta and
Monte-Carlo). They share the same point estimate but differ in their standard
errors and intervals. By default, `coef()` and `confint()` use the method that
also underlies the Zhao, Lynch & Chen approach (Monte-Carlo for `lavaan` and
`modsem`, bootstrap for `cSEM`, and the posterior for `blavaan`). Use the
`method` argument to choose a different one:

```{r}
confint(out, parm = "indirect", method = "sobel")
```

The level of all intervals is set when calling `rmedsem()` with the
`ci.two.tailed` argument:

```{r}
out90 <- rmedsem(mod, indep = "math", med = "read", dep = "science",
                 ci.two.tailed = 0.90)
confint(out90)
```

`as.data.frame()` returns the estimates of the indirect effect for all methods
as a data frame:

```{r}
as.data.frame(out)
```

## Effect sizes

The effect sizes are available through dedicated functions:

```{r}
RIT(out)      # ratio of indirect to total effect
RID(out)      # ratio of indirect to direct effect
Upsilon(out)  # variance in 'science' explained indirectly by 'math' via 'read'
Upsilon(out, adjusted = FALSE)
```

## Plots

`plot()` shows a coefficient plot of the indirect (for each method), direct
and total effects, or a pie chart of the indirect and direct effects:

```{r, fig.width=7, fig.height=4}
plot(out)
```

```{r, fig.width=6, fig.height=6}
plot(out, type = "effect")
```

## Bayesian models: equal-tailed intervals or HDI

For models estimated with `blavaan`, the intervals are credible intervals
computed from the posterior samples. By default, these are equal-tailed
intervals (the 2.5% and 97.5% quantiles for `ci.two.tailed = 0.95`). Set
`hdi = TRUE` to obtain highest density intervals (HDI) instead, which requires
the [HDInterval](https://cran.r-project.org/package=HDInterval) package:

```{r, include = FALSE}
has.blavaan <- requireNamespace("blavaan", quietly = TRUE) &&
  requireNamespace("HDInterval", quietly = TRUE)
```

```{r, eval = has.blavaan, results = "hide"}
library(blavaan)
bmod <- bsem(mod.txt, data = hsbdemo, n.chains = 3, burnin = 500, sample = 500,
             seed = 1, bcontrol = list(cores = 3, refresh = 0))
```

```{r, eval = has.blavaan}
out.ci  <- rmedsem(bmod, indep = "math", med = "read", dep = "science")
out.hdi <- rmedsem(bmod, indep = "math", med = "read", dep = "science",
                   hdi = TRUE)

confint(out.ci)
confint(out.hdi)
summary(out.hdi)
```
